---
title: "SQANTI-single cell<br>multi-sample `r if (exists('mode') && mode == 'isoforms') 'isoforms' else 'reads'` report"
date: "`r paste0('Date: ', format(Sys.time(), '%d %B %Y'))`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
    number_sections: false
    css: style-multisample.css
    theme:
      version: 3
      bootswatch: flatly
    highlight: tango
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(dplyr)
library(htmltools)
```

# Samples Per-Cell Summary

```{r summary-table}
if (exists("multi_summary_tbl_html", envir = .GlobalEnv)) {
  summary_tbl <- get("multi_summary_tbl_html", envir = .GlobalEnv)
  knitr::kable(summary_tbl, format = "html", escape = FALSE, align = "c")
} else {
  cat("No Samples Per-Cell Summary table available.")
}
```

# Samples Per-Cell Library Size

```{r libsize-dropdown}
if (exists("multi_library_size_plots", envir = .GlobalEnv)) {
  lplots <- get("multi_library_size_plots", envir = .GlobalEnv)
  lib_names <- names(lplots)

  if (length(lib_names) > 0) {
    items_html <- paste(sapply(lib_names, function(name) {
      target_id <- paste0("libsize-", gsub("[^A-Za-z0-9]", "-", tolower(name)))
      sprintf('      <li><a href="#" data-target="#%s">%s</a></li>', target_id, htmltools::htmlEscape(name))
    }), collapse = "\n")

    htmltools::HTML(paste0(
      '<div id="libsize-dropdown" class="dropdown-toolbar">\n',
      '  <div class="btn-group">\n',
      '    <button type="button" class="btn btn-default dropdown-toggle" data-toggle="dropdown">\n',
      '      <span class="label">', htmltools::htmlEscape(lib_names[1]), '</span> <span class="caret"></span>\n',
      '    </button>\n',
      '    <ul class="dropdown-menu dropdown-menu-right" role="menu">\n',
      items_html, '\n',
      '    </ul>\n',
      '  </div>\n',
      '</div>\n'
    ))
  }
} else {
  cat("No per-cell library size plots available.")
}
```

```{r libsize-panes, results='asis'}
if (exists("multi_library_size_plots", envir = .GlobalEnv)) {
  lplots <- get("multi_library_size_plots", envir = .GlobalEnv)
  first <- TRUE
  for (i in seq_along(names(lplots))) {
    plot_name <- names(lplots)[i]
    target_id <- paste0("libsize-", gsub("[^A-Za-z0-9]", "-", tolower(plot_name)))
    css_class <- if (first) "toggle-pane" else "toggle-pane hidden-pane"
    css_style <- if (first) "display:block;" else "display:none;"
    first <- FALSE

    res <- knitr::knit_child(text = c(
      sprintf('<div id="%s" class="%s" style="%s">', target_id, css_class, css_style),
      '',
      sprintf('```{r libsize-plot-%d, fig.width=12, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}', i),
      'lp <- get("multi_library_size_plots", envir = .GlobalEnv)',
      sprintf('print(lp[["%s"]])', plot_name),
      '```',
      '',
      '</div>'
    ), envir = globalenv(), quiet = TRUE)
    cat(res, sep = '\n')
  }
}
```

```{=html}
<div class="section-spacer-xl"></div>
```

# Samples Gene Characterization

```{r genechar-dropdown}
if (exists("multi_gene_char_plots", envir = .GlobalEnv)) {
  gplots <- get("multi_gene_char_plots", envir = .GlobalEnv)
  gene_names <- names(gplots)

  if (length(gene_names) > 0) {
    items_html <- paste(sapply(gene_names, function(name) {
      target_id <- paste0("genechar-", gsub("[^A-Za-z0-9]", "-", tolower(name)))
      sprintf('      <li><a href="#" data-target="#%s">%s</a></li>', target_id, htmltools::htmlEscape(name))
    }), collapse = "\n")

    htmltools::HTML(paste0(
      '<div id="genechar-dropdown" class="dropdown-toolbar">\n',
      '  <div class="btn-group">\n',
      '    <button type="button" class="btn btn-default dropdown-toggle" data-toggle="dropdown">\n',
      '      <span class="label">', htmltools::htmlEscape(gene_names[1]), '</span> <span class="caret"></span>\n',
      '    </button>\n',
      '    <ul class="dropdown-menu dropdown-menu-right" role="menu">\n',
      items_html, '\n',
      '    </ul>\n',
      '  </div>\n',
      '</div>\n'
    ))
  }
} else {
  cat("No gene characterization plots available.")
}
```

```{r genechar-panes, results='asis'}
if (exists("multi_gene_char_plots", envir = .GlobalEnv)) {
  gplots <- get("multi_gene_char_plots", envir = .GlobalEnv)
  first <- TRUE
  for (i in seq_along(names(gplots))) {
    plot_name <- names(gplots)[i]
    target_id <- paste0("genechar-", gsub("[^A-Za-z0-9]", "-", tolower(plot_name)))
    css_class <- if (first) "toggle-pane" else "toggle-pane hidden-pane"
    css_style <- if (first) "display:block;" else "display:none;"
    first <- FALSE

    res <- knitr::knit_child(text = c(
      sprintf('<div id="%s" class="%s" style="%s">', target_id, css_class, css_style),
      '',
      sprintf('```{r genechar-plot-%d, fig.width=12, fig.height=7, echo=FALSE, warning=FALSE, message=FALSE}', i),
      'gp <- get("multi_gene_char_plots", envir = .GlobalEnv)',
      sprintf('print(gp[["%s"]])', plot_name),
      '```',
      '',
      '</div>'
    ), envir = globalenv(), quiet = TRUE)
    cat(res, sep = '\n')
  }
}
```

```{=html}
<div class="section-spacer-xl"></div>
```

# Samples `r entity_label_plural` Length Distribution

```{r length-dist-plot, fig.width=12, fig.height=7}
if (exists("multi_length_distribution_plot", envir = .GlobalEnv)) {
  print(get("multi_length_distribution_plot", envir = .GlobalEnv))
} else {
  cat("Length distribution plot is not available (classification files were not provided).")
}
```

```{=html}
<div class="section-spacer-xl"></div>
```

# Samples `r entity_label_plural` Distributions per Structural Category {.tabset .tabset-fade}

## Samples Comparison

```{=html}
<div id="structcat-dropdown" class="dropdown-toolbar">
  <div class="btn-group">
    <button type="button" class="btn btn-default dropdown-toggle" data-toggle="dropdown">
      <span class="label">FSM</span> <span class="caret"></span>
    </button>
    <ul class="dropdown-menu dropdown-menu-right" role="menu">
      <li><a href="#" data-target="#structcat-fsm">FSM</a></li>
      <li><a href="#" data-target="#structcat-ism">ISM</a></li>
      <li><a href="#" data-target="#structcat-nic">NIC</a></li>
      <li><a href="#" data-target="#structcat-nnc">NNC</a></li>
      <li><a href="#" data-target="#structcat-genic-genomic">Genic Genomic</a></li>
      <li><a href="#" data-target="#structcat-antisense">Antisense</a></li>
      <li><a href="#" data-target="#structcat-fusion">Fusion</a></li>
      <li><a href="#" data-target="#structcat-intergenic">Intergenic</a></li>
      <li><a href="#" data-target="#structcat-genic-intron">Genic Intron</a></li>
    </ul>
  </div>
</div>
```

<div id="structcat-fsm" class="toggle-pane" style="display:block;">

```{r structcat-fsm, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["FSM"]])) print(plots[["FSM"]])
}
```

</div>

<div id="structcat-ism" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-ism, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["ISM"]])) print(plots[["ISM"]])
}
```

</div>

<div id="structcat-nic" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-nic, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["NIC"]])) print(plots[["NIC"]])
}
```

</div>

<div id="structcat-nnc" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-nnc, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["NNC"]])) print(plots[["NNC"]])
}
```

</div>

<div id="structcat-genic-genomic" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-genic-genomic, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["Genic Genomic"]])) print(plots[["Genic Genomic"]])
}
```

</div>

<div id="structcat-antisense" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-antisense, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["Antisense"]])) print(plots[["Antisense"]])
}
```

</div>

<div id="structcat-fusion" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-fusion, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["Fusion"]])) print(plots[["Fusion"]])
}
```

</div>

<div id="structcat-intergenic" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-intergenic, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["Intergenic"]])) print(plots[["Intergenic"]])
}
```

</div>

<div id="structcat-genic-intron" class="toggle-pane hidden-pane" style="display:none;">

```{r structcat-genic-intron, fig.width=12, fig.height=7}
if (exists("multi_structural_category_violin_plots", envir = .GlobalEnv)) {
  plots <- get("multi_structural_category_violin_plots", envir = .GlobalEnv)
  if (!is.null(plots[["Genic Intron"]])) print(plots[["Genic Intron"]])
}
```

</div>

```{=html}
<div class="section-spacer-xl"></div>
```

# PCA Analysis {.tabset .tabset-fade}

## PCA Plot

```{r pca-scores-dropdown}
if (exists("multi_pca_group_plot", envir = .GlobalEnv)) {
  htmltools::HTML(paste0(
    '<div id="pca-sample-group-dropdown" class="dropdown-toolbar">\n',
    '  <div class="btn-group">\n',
    '    <button type="button" class="btn btn-default dropdown-toggle" data-toggle="dropdown">\n',
    '      <span class="label">By Sample</span> <span class="caret"></span>\n',
    '    </button>\n',
    '    <ul class="dropdown-menu dropdown-menu-right" role="menu">\n',
    '      <li><a href="#" data-target="#pca-plot-by-sample">By Sample</a></li>\n',
    '      <li><a href="#" data-target="#pca-plot-by-group">By Group</a></li>\n',
    '    </ul>\n',
    '  </div>\n',
    '</div>\n'
  ))
}
```

```{r pca-plot-panes, results='asis'}
if (exists("multi_pca_group_plot", envir = .GlobalEnv)) {
  res1 <- knitr::knit_child(text = c(
    '<div id="pca-plot-by-sample" class="toggle-pane" style="display:block;">',
    '',
    '```{r pca-by-sample-inner, fig.width=12, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}',
    'if (exists("multi_pca_scores_plot", envir = .GlobalEnv)) {',
    '  print(get("multi_pca_scores_plot", envir = .GlobalEnv))',
    '} else { cat("PCA score plot is not available.") }',
    '```',
    '',
    '</div>'
  ), envir = globalenv(), quiet = TRUE)
  cat(res1, sep = '\n')

  res2 <- knitr::knit_child(text = c(
    '<div id="pca-plot-by-group" class="toggle-pane hidden-pane" style="display:none;">',
    '',
    '```{r pca-by-group-inner, fig.width=12, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}',
    'print(get("multi_pca_group_plot", envir = .GlobalEnv))',
    '```',
    '',
    '</div>'
  ), envir = globalenv(), quiet = TRUE)
  cat(res2, sep = '\n')
} else {
  res <- knitr::knit_child(text = c(
    '```{r pca-scores-plain, fig.width=12, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}',
    'if (exists("multi_pca_scores_plot", envir = .GlobalEnv)) {',
    '  print(get("multi_pca_scores_plot", envir = .GlobalEnv))',
    '} else { cat("PCA score plot is not available.") }',
    '```'
  ), envir = globalenv(), quiet = TRUE)
  cat(res, sep = '\n')
}
```

## Scree Plot

<div id="pca-scree" class="pca-pane">

```{r pca-scree, fig.width=12, fig.height=8}
if (exists("multi_pca_scree_plot", envir = .GlobalEnv)) {
  print(get("multi_pca_scree_plot", envir = .GlobalEnv))
} else {
  cat("PCA scree plot is not available.")
}
```

</div>

## Top 10 loadings by component {.tabset}

### PC Loadings

```{r pcaload-bar-dropdown}
if (exists("multi_pca_top_loadings_plots", envir = .GlobalEnv)) {
  lplots <- get("multi_pca_top_loadings_plots", envir = .GlobalEnv)
  pc_names <- paste0(names(lplots), " Loadings")

  items_html <- paste(sapply(pc_names, function(name) {
    target_id <- paste0("pcaload-", gsub("[^A-Za-z0-9]", "-", tolower(name)))
    sprintf('      <li><a href="#" data-target="#%s">%s</a></li>', target_id, htmltools::htmlEscape(name))
  }), collapse = "\n")

  htmltools::HTML(paste0(
    '<div id="pcaload-dropdown" class="dropdown-toolbar">\n',
    '  <div class="btn-group">\n',
    '    <button type="button" class="btn btn-default dropdown-toggle" data-toggle="dropdown">\n',
    '      <span class="label">', htmltools::htmlEscape(pc_names[1]), '</span> <span class="caret"></span>\n',
    '    </button>\n',
    '    <ul class="dropdown-menu dropdown-menu-right" role="menu">\n',
    items_html, '\n',
    '    </ul>\n',
    '  </div>\n',
    '</div>\n'
  ))
}
```

```{r pcaload-bar-panes, results='asis'}
if (exists("multi_pca_top_loadings_plots", envir = .GlobalEnv)) {
  lplots <- get("multi_pca_top_loadings_plots", envir = .GlobalEnv)
  first <- TRUE
  for (i in seq_along(names(lplots))) {
    pc_name <- names(lplots)[i]
    display_name <- paste0(pc_name, " Loadings")
    target_id <- paste0("pcaload-", gsub("[^A-Za-z0-9]", "-", tolower(display_name)))
    css_class <- if (first) "toggle-pane" else "toggle-pane hidden-pane"
    css_style <- if (first) "display:block;" else "display:none;"
    first <- FALSE

    res <- knitr::knit_child(text = c(
      sprintf('<div id="%s" class="%s" style="%s">', target_id, css_class, css_style),
      '',
      sprintf('```{r pcaload-bar-%d, fig.width=12, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}', i),
      sprintf('lp <- get("multi_pca_top_loadings_plots", envir = .GlobalEnv)'),
      sprintf('print(lp[["%s"]])', pc_name),
      '```',
      '',
      '</div>'
    ), envir = globalenv(), quiet = TRUE)
    cat(res, sep = '\n')
  }
}
```

### Feature Distributions

```{r pcafeat-dist-dropdown}
if (exists("multi_pca_loading_distribution_plots", envir = .GlobalEnv)) {
  fplots <- get("multi_pca_loading_distribution_plots", envir = .GlobalEnv)
  feat_names <- names(fplots)

  if (length(feat_names) > 0) {
    items_html <- paste(sapply(feat_names, function(name) {
      target_id <- paste0("pcafeat-", gsub("[^A-Za-z0-9]", "-", tolower(name)))
      sprintf('      <li><a href="#" data-target="#%s">%s</a></li>', target_id, htmltools::htmlEscape(name))
    }), collapse = "\n")

    htmltools::HTML(paste0(
      '<div id="pcafeat-dropdown" class="dropdown-toolbar">\n',
      '  <div class="btn-group">\n',
      '    <button type="button" class="btn btn-default dropdown-toggle" data-toggle="dropdown">\n',
      '      <span class="label">', htmltools::htmlEscape(feat_names[1]), '</span> <span class="caret"></span>\n',
      '    </button>\n',
      '    <ul class="dropdown-menu dropdown-menu-right" role="menu">\n',
      items_html, '\n',
      '    </ul>\n',
      '  </div>\n',
      '</div>\n'
    ))
  }
} else {
  cat("No feature distribution plots available.")
}
```

```{r pcafeat-dist-panes, results='asis'}
if (exists("multi_pca_loading_distribution_plots", envir = .GlobalEnv)) {
  fplots <- get("multi_pca_loading_distribution_plots", envir = .GlobalEnv)
  first <- TRUE
  for (i in seq_along(names(fplots))) {
    feat_name <- names(fplots)[i]
    target_id <- paste0("pcafeat-", gsub("[^A-Za-z0-9]", "-", tolower(feat_name)))
    css_class <- if (first) "toggle-pane" else "toggle-pane hidden-pane"
    css_style <- if (first) "display:block;" else "display:none;"
    first <- FALSE

    res <- knitr::knit_child(text = c(
      sprintf('<div id="%s" class="%s" style="%s">', target_id, css_class, css_style),
      '',
      sprintf('```{r pcafeat-dist-%d, fig.width=12, fig.height=8, echo=FALSE, warning=FALSE, message=FALSE}', i),
      sprintf('fp <- get("multi_pca_loading_distribution_plots", envir = .GlobalEnv)'),
      sprintf('print(fp[["%s"]])', feat_name),
      '```',
      '',
      '</div>'
    ), envir = globalenv(), quiet = TRUE)
    cat(res, sep = '\n')
  }
}
```

```{=html}
<script>
(function(){
  function setupDropdown(groupId){
    var container=document.getElementById(groupId);
    if(!container) return;
    var labelSpan=container.querySelector(".dropdown-toggle .label");
    var links=container.querySelectorAll(".dropdown-menu a");
    var panesContainer = container.parentElement;
    while (panesContainer && !panesContainer.querySelector('.toggle-pane')) {
      panesContainer = panesContainer.parentElement;
    }
    if (!panesContainer) panesContainer = container.closest('section, div.section') || document;

    Array.prototype.forEach.call(links,function(link){
      link.addEventListener("click",function(e){
        e.preventDefault();
        var target = panesContainer.querySelector(this.getAttribute("data-target"));
        if (!target) return;
        Array.prototype.forEach.call(panesContainer.querySelectorAll(".toggle-pane"), function(el){
          el.classList.add("hidden-pane");
          el.style.display = "none";
        });
        target.classList.remove("hidden-pane");
        target.style.display = "block";
        if(labelSpan) labelSpan.textContent=this.textContent.trim();
        setTimeout(function(){
          if (window.jQuery && jQuery.fn.dataTable) {
            jQuery(target).find('table.dataTable').each(function(){
              if (jQuery.fn.dataTable.isDataTable(this)) {
                jQuery(this).DataTable().columns.adjust().draw(false);
              }
            });
          }
        }, 0);
      });
    });
  }

  document.addEventListener("DOMContentLoaded",function(){
    setupDropdown("libsize-dropdown");
    setupDropdown("genechar-dropdown");
    setupDropdown("structcat-dropdown");
    setupDropdown("pca-sample-group-dropdown");
    setupDropdown("pcaload-dropdown");
    setupDropdown("pcafeat-dropdown");
  });
})();
</script>
```
